### 2020-10-08 response.tex

Done:

``For example, in our analysis of the COVID-19 serial intervals from China, outside Hubei province, we find that the original \Ro estimates based on aggregated serial-interval data underestimated \Ro by a factor of 2.0--2.6.''

\revtext{Figure 1 clearly illustrates the different intervals from different perspectives but does not explain the source of biases. Consider ways to visually communicate the important insights at the end of section 3.1.}

Thank you for the suggestion, although we are not positive we understand this comment. Figure 1 is not really intended to communicate the source of biases: that comes later in the course of our exposition. We have made the following changes to attempt to improve our visual communication.

@@

We have added some visual aids to Figure 1: we color intrinsic, backward, and forward intervals and explain them briefly in the caption:

``
Intrinsic intervals (black) reflect average of individual characteristics and are not dependent on population-level dynamics.
  Forward intervals (green) can change due to epidemiological dynamics (e.g., contraction of generation intervals through susceptible depletion).
  Backward intervals (blue) can change due to changes in cohort sizes even when forward intervals remain time-invariant.
''

We also color forward and backward intervals in Figure 3 according to the color scheme in Figure 1.
We previously referred to figure 1 at the end of Section 3.1 (now Section 3.2); this paragraph (quoted here) carefully outlines the dynamical effects on the realized lengths of the different intervals.

@@

``The forward serial-interval distributions depend on distributions of three intervals
(Fig. 1{diagram}B): (i) the backward incubation period, (ii) the forward generation interval, and (iii) the forward incubation period.
In these simulations, both forward incubation period (Fig. 3B) and generation-interval (Fig. 3C) distributions remain roughly constant;
therefore, changes in the forward serial-interval distributions (Fig. 3D) are predominantly driven by changes in the backward incubation period distribution, whose mean increases over time as the growth rate of disease incidence slows and then reverses.
In general, relative contributions of the three distributions depend on their shapes, correlations between intrinsic incubation periods and generation intervals, and overall epidemiological dynamics.''

We are concerned that making further changes to the figures would probably make them overly complex, and therefore harder to interpret.

### 2020-10-08 serial.tex
abstract tweaks

The reproduction number \RR and the exponential growth rate $r$ are critical quantities for characterizing outbreak dynamics.
These quantities are linked by generation intervals---defined as the time between infection and transmission.
Because generation intervals are difficult to measure directly, serial intervals---defined as the onset of symptoms in a chain of transmission---are often substituted.
However, recent studies suggest that the two intervals give different estimates of \RR from $r$.
We show that \emph{forward-looking} serial intervals, which begin in a certain period, correctly describe the renewal process of symptomatic cases and therefore link \RR with $r$.
These forward intervals are different from \emph{intrinsic} intervals, which neglect population-level dynamics, and \emph{backward-looking} intervals, which end in a certain period.
Instead, the forward-looking intervals depend on $r$:
When an epidemic is growing exponentially, infectors that develop symptoms on a given day will have shorter incubation periods, on average, than their infectees, thereby affecting the serial-interval distribution.
We provide a heuristic method for addressing biases that can arise from neglecting changes in serial intervals. 
We apply the method to estimation of \RR for the COVID-19 outbreak in China, outside Hubei province, using serial-interval data collected between January 21--February 8, 2020. 
Our analysis shows that using incorrectly defined serial intervals can bias estimates of initial \RR by up to a factor of 2.6 in the context of COVID-19.
This study demonstrates the importance of early contact tracing efforts and provides a rationale for reassessing generation intervals, serial intervals, and \RR\ estimates, for COVID-19.

>>>>

The reproduction number \RR and the exponential growth rate $r$ are critical quantities for characterizing outbreak dynamics.
These quantities are linked by generation intervals, defined as the time between infection and transmission.
Because generation intervals are difficult to measure directly, serial intervals---defined as the onset of symptoms in a chain of transmission---are often substituted.
However, recent studies suggest that making this substitution can lead to significant changes when using $r$ to estimate \RR.
Here we explore the differences among these intervals, how they vary over the course of an epidemic, and the implications for estimating epidemic parameters.
We show that \emph{forward-looking} serial intervals, which begin in a certain period, correctly describe the renewal process of symptomatic cases and therefore reliably link \RR with $r$.
These forward intervals are different from \emph{intrinsic} intervals, which neglect population-level dynamics, and \emph{backward-looking} intervals, which end in a certain period.
The forward-looking intervals depend on $r$:
During the exponential growth phase of an epidemic, infectors that develop symptoms on a given day will have shorter incubation periods, on average, than their infectees, changing the serial-interval distribution in complex ways that depend on 
We provide a heuristic method for addressing biases that can arise from neglecting changes in serial intervals. 
We apply the method to estimation of \RR for the COVID-19 outbreak in China, outside Hubei province, using serial-interval data collected between January 21--February 8, 2020. 
Our analysis shows that using improperly defined serial intervals can bias estimates of initial \RR by up to a factor of 2.6 in the context of COVID-19.
This study demonstrates the importance of early contact tracing efforts and provides a rationale for reassessing generation intervals, serial intervals, and \RR\ estimates, for COVID-19.
